TacticToe: Learning to Prove with Tactics

### 2018-02-20 tactictoe_jv.tex
minor
introduction of evaluation

\section{Search tree}\label{sec:prelim}

The application of a tactic to an initial conjecture produces a list of goals.
Each of the created goals can in turn be the input to other tactics.
Moreover, it is possible to try multiple tactics on the same goal.
In order to keep track of the progress made so far from the initial conjecture we want to solve, we organize goals and tactics into a graph with list of goals as nodes and tactics as edges (See Definition~\ref{def:stree}).

@@

\paragraph{Initialization}
The input of the algorithm is a goal $g$ (also called conjecture) that we
want to prove.
\todo{@TG: Check sentence below}
Therefore the search tree starts with only one node the list of goals
$[g]$ and an ordered list of predicted tactics for the goal $g$.
%Then, the MCTS algorithm alternates node selection steps and node extension
%steps which grow the search tree.

@@

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e., the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e., there are no tactics to be applied to any open goal.
This happens less than 10 times in the full-scale experiment. \todo{@TG: why don't we have the original tactics as options?}
\item It exceeds a given time limit (default is set to 60 seconds).
\end{itemize}

@@

\section{Feature Extraction}\label{sec:features}
\todo{@TG: decide which of the next sentences}
We explain the construction of the similarity measure by which relevant
objects are predicted

We predicts tactics through their associated goals. So, the features
for each kind of object can be extracted from (different representations of)
mathematical formulas. We start by describing features for \holfour terms. From
there, we extend the extraction mechanism
to goals, theorems and lists of goals. Duplicated features are always removed
so that each object has an associated set of features.

@@

To preselect a list of goal $l$ from our database of lists of goals, we
consider the tactic input $g$ that was at the origin of the production of $l$
during orthogonalization. The list $l$
is preselected if and only if $g$ appears in the 500 preselected tactic-goal
pairs. This has the
desirable effect that negative and positive examples derived from the same
input goal are chosen or discarded together.
\todo{this paragraph was confusing for Lasse and Cezary}

@@

Fix parameters: not optimized here.
 - Features extraction method.
 - k-nn distance.
 - MCTS evaluation radius is set to 10.
 - Metis : 16 lemmas.
 - Eprover: 128 lemmas and 5 seconds timeout 1 asynchronous call

@@

In Table~\ref{tab:tuning}, we run \tactictoe with different parameters.
The first four techniques are tested relative to the same baseline indicated in
the table by the tag ``default''.  This default strategy relies on a tactic
time out of 0.05s, an orthogonalization radius of 20 and a policy coefficent of
0.5.
The two last techniques are tested relative
to a baseline consisting of the best set of parameters discovered so far and
are marked with the improvement over the ``default'' strategy.
In each subtable, each experiment differs from the current baseline by exactly
one parameter.

@@

@@

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
generate
proof scripts from previous \coq proof examples.
Its strength lies in the ability to produce likely sequences
of tactics for solving domain specific goals. It operates by creating a model
for common sequences of tactics for a specific library.
This means that in order to propose the next tactic, only the previously
called tactics
are considered.
%intuition tells us that it should not be the main factor
Our algorithm, on the other hand, relies mainly on the characteristics of the
current goal
to decide
which tactics to apply next. In this way, our learning mechanism has to
rediscover why each
tactic was applied for the current subgoals. It may lack some useful bias for
common sequences
of tactics, but is more reactive to subtle changes. Indeed, it can be trained
on a large library and only tactics relevant to the current subgoal will be
selected.
Concerning the proof search, \textsf{SEPIA}'s %simple
breadth-first search is replaced by MCTS which allows for supervised learning
guidance in the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts of the
\coq library demonstrating that it globally outperforms individual \coq
tactics. Here, we demonstrate the competitiveness of our system against
\eprover on the \holfour standard library.
%A recent improvement on this system~\cite{}\todo{Cezary: could not remember
%the
%reference that you gave me here i think it was CICM 2017} includes the
%possibility to
%substitute theorems appearing in tactics by ones belonging to the same
%cluster.
%In this work, we are also able to predict theorems appearing as arguments of
%tactics but we rely only on information from the current goal. Combining the
%two approaches by predicting from the goal and from the original tactic
%arguments is more difficult but is likely to produce more suitable lemmas.
%Proof patching: \cite{RingerYLG18}

\todoi{maybe the paragraph can be generally proof strategy languages including ltac and eisbach}
A similar effort for stronger automation is performed in \isabelle. Nagashima
developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
user to
program with regular tactics and combine them with tools such as
\sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}.
Selection of tactics from data collected in recorded proofs is in development
but for now it is up to the user to come up with a good meta-tactic.

>>>>

\section{Search Tree}\label{sec:prelim}
The application of a tactic to an initial conjecture produces a list of goals.
Each of the created goals can in turn be the input to other tactics.
Moreover, it is possible to try multiple tactics on the same goal.
In order to keep track of the progress made so far from the initial conjecture 
we want to solve, we organize goals and tactics into a graph with list of goals 
as nodes and tactics as edges (See Definition~\ref{def:stree}).

@@

\paragraph{Initialization}
The input of the algorithm is a goal $g$ (also called conjecture) that we
want to prove. Therefore the search tree starts with only one node containing 
the list of goals
$[g]$.
%Then, the MCTS algorithm alternates node selection steps and node extension
%steps which grow the search tree.

@@

\paragraph{Resolution}
The search ends when the algorithm reaches one of these 3 conditions:
\begin{itemize}
\item It finds a proof, i.e., the root node is solved. In this case,
the search returns a minimized and
prettified proof script (see Section~\ref{sec:proofdisplay}).
\item It saturates, i.e., there are no tactics to be applied to any open goal.
This occurs less than 10 times in the full-scale experiment.
This happens only at the beginning of \holfour library, where the training data 
has very few tactics. 
\item It exceeds a given time limit (default is set to 60 seconds).
\end{itemize}

@@

\section{Feature Extraction}\label{sec:features}
%We explain the construction of the similarity measure by which relevant
%objects are predicted

We predict tactics through their associated goals. So, the features
for each kind of object can be extracted from (different representations of)
mathematical formulas. We start by describing features for \holfour terms. From
there, we extend the extraction mechanism
to goals, theorems and lists of goals. Duplicated features are always removed
so that each object has an associated set of features.

@@

We do preselection also for lists of goals.
To preselect a list of goals $l$ from our database of lists of goals, we
consider the tactic input $g$ that was at the origin of the production of $l$
during orthogonalization. The list $l$ is preselected if and only if $g$ 
appears in the 500 preselected tactic-goal pairs. 
%This has the
%desirable effect that negative and positive examples derived from the same
%input goal are chosen or discarded together.

@@

We optimize \tactictoe by tuning parameters of six different techniques:
the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS 
evaluation and ATP integration. During training, optional techniques such as 
orthogonalization may also be turned off completely.
\tactictoe also includes other important techniques which are run with their 
default parameters: features extraction, prediction 
algorithms, number of premises for ATPs and MCTS evaluation radius (set to 
10).
By choosing a set of parameters, we create a strategy for 
\tactictoe that is evaluated on our training set of 273 theorems. And the 
strategy with the highest number of proven theorems is selected for 
a full scale evaluation.

@@

In Table~\ref{tab:tuning}, the success rate of the different \tactictoe  
strategy is presented.
The first four techniques are tested relative to the same baseline indicated in
the table by the tag ``default''.  This default strategy relies on a tactic
time out of 0.05s, an orthogonalization radius of 20 and a policy coefficent of
0.5.
The two last techniques are tested relative
to a baseline consisting of the best set of parameters discovered so far and
are marked with the improvement over the ``default'' strategy.
In each subtable, each experiment differs from the current baseline by exactly
one parameter.

@@

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
generate
proof scripts from previous \coq proof examples.
Its strength lies in the ability to produce likely sequences
of tactics for solving domain specific goals. It operates by creating a model
for common sequences of tactics for a specific library.
This means that in order to propose the next tactic, only the previously
called tactics
are considered.
%intuition tells us that it should not be the main factor
Our algorithm, on the other hand, relies mainly on the characteristics of the
current goal
to decide
which tactics to apply next. In this way, our learning mechanism has to
rediscover why each
tactic was applied for the current subgoals. It may lack some useful bias for
common sequences
of tactics, but is more reactive to subtle changes. Indeed, it can be trained
on a large library and only tactics relevant to the current subgoal will be
selected.
Concerning the proof search, \textsf{SEPIA}'s %simple
breadth-first search is replaced by MCTS which allows for supervised learning
guidance in the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts of the
\coq library demonstrating that it globally outperforms individual \coq
tactics. Here, we demonstrate the competitiveness of our system against
\eprover on the \holfour standard library.

@@

%A recent improvement on this system~\cite{}\todo{Cezary: could not remember
%the
%reference that you gave me here i think it was CICM 2017} includes the
%possibility to
%substitute theorems appearing in tactics by ones belonging to the same
%cluster.
%In this work, we are also able to predict theorems appearing as arguments of
%tactics but we rely only on information from the current goal. Combining the
%two approaches by predicting from the goal and from the original tactic
%arguments is more difficult but is likely to produce more suitable lemmas.
%
Proof patching: \cite{RingerYLG18}

\todoi{maybe the paragraph can be generally proof strategy languages including ltac and eisbach}
A similar effort for stronger automation is performed in \isabelle. Nagashima
developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
user to
program with regular tactics and combine them with tools such as
\sledgehammer.
Selection of tactics from data collected in recorded proofs is in development
but for now it is up to the user to come up with a good meta-tactic.
